---
title: "Sources of nutrients, their variability and contribution to the total input into aquaponic systems"
author:
  - name: Anıl Axel Tellbüscher
    email: atellbuscher@frov.jcu.cz
    affiliation: USB
    footnote: 1
  - name: Radek Gebauer
    email: rgebauer@frov.jcu.cz
    affiliation: USB
    footnote: 2
  - name: Jan Mráz
    email: jmraz@frov.jcu.cz
    affiliation: USB
    correspondingauthor: true
address:
  - code: USB
    organization: University of South Bohemia in České Budějovice, Faculty of Fisheries and Protection of Waters, South Bohemian Research Center of Aquaculture and Biodiversity of Hydrocenoses, Institute of Aquaculture and Protection of Waters
    addressline: Na Sádkach 1780
    city: 37005 České Budějovice
    country: Czech Republic
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Input routes of nutrients into aquaculture and aquaponic systems are known, but their contribution to the total nutrient input and the variability between locations remains unclear.
keywords: 
  - aquaponics
  - minerals
  - fish feed
  - alkalinity supplements
journal: "Reviews in Aquaculture"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: ../references.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    includes:
      in_header: ../latex/preamble.tex
    keep_tex: true
    citation_package: default
---

```{r setup, include=FALSE}
# Knitr options
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
options(digits = 1)

# Load packages
library(here)
library(readxl)
library(tidyverse)
library(ggmap)
library(gdata)
library(viridis)
library(RColorBrewer)
library(bookdown)
library(png)
library(tellbuschR) # Install from GitHub TellAnAx/tellbuschR with GitHub token for authentication

# Load functions
source(here("R", "plot_functions.R"))
```

# Introduction
One of the obstacles in the wider adoption of aquaponic systems, the combination of aquaculture and hydroponics (soilless plant production), is stated to be the rather complex management of plant nutrients within the system. While a wide range of "off-the-shelf" fertilizer is available for hydroponics, this is not the case for aquaponic systems. Developing more efficient and simple nutrient management practices thus appears to be an opportunity to facilitate systems management. Within this context, nutrient management refers to guaranteeing that optimum plant nutrient concentrations are met by controlling the inputs.

In aquaculture and aquaponics, nutrients enter the system through the daily ration of feed for the aquatic livestock (hereinafter denoted as aquafeed), alkalinity supplements used for pH control, and the source water [@Rakocy2006; @Eck2019; @Robaina2019]. Within this context, essential plant nutrients are the elements nitrogen, phosphorus, potassium, calcium, magnesium, sulfur, iron, boron, silicon, manganese, copper, zinc, nickel, and molybdenum [Mengel].

When focussing on the efficiency aspect of nutrient management, the most efficient solution would be to monitor each plant nutrient individually. This would ensure that each individual nutrient can be supplied in the necessary amount to reach target concentrations, for instance as pure salts. However, a full-spectrum monitoring of all plant nutrients is not practical, as it requires expensive equipment and qualified workforce. Instead, it would be more meaningful to focus on the simplicity aspect and develop a management practice that is easily implementable without introducing additional tasks to the working schedule. Against this background, optimizing an existing nutrient input can be considered as more feasible solution.

An approach to optimize nutrient inputs that is followed is the development of aquafeeds that are formulated in a way that the excreta of the livestock meet plant nutrient demands. Though, when optimizing one nutrient input while not controlling for the others, it is important to consider the range of nutrient input that can be expected by other sources of nutrients. Awareness about the nutrients that run the risk of under- or oversupply under "non-average" conditions is crucial, as this might have adverse effects on the livestock or plants. Hence, the variability of nutrients originating from different sources ultimately determines which nutrients to focus on when optimizing nutrient inputs. However,little is known about the variability of plant nutrient mass flows in aquaculture, which is crucial for the development of a nutrient management solution that is optimized with regards to its simplicity of use, such as a dedicated aquafeed. The objective of this study is thus to assess the nutrient budget of aquaculture systems with regards to the contribution and variability of aquafeeds, alkalinity supplements, and the source water. Special emphasis is given to aquaponics as this knowledge is expected to contribute to the development of simple means of nutrient management in this food production system.





# Methodology

## Model system

```{r}
source(here("R", "assumptions.R"))
```

The inputs of aquafeeds, alkalinity supplements and source water was based on an exemplary \gls{ras}: The ASTAF-PRO system [@Kloas2015] was chosen as model because its sizing is close to reality when comparing it with results from a recent survey targeting aquaponics production sites [@Pattillo2020] and its system components are well-described. All model data are summarized in table \ref{tab:assumptions}. 

\include{table_assumptions.tex}



## Data collection and calculations

### Source water

```{r import_data}
source(here("R", "import_data.R"))
```

```{r data_wrangling}
source(here("R", "data_wrangling.R"))
```

```{r plot_functions}
source(here("R", "plot_functions.R"))
```

```{r plot_water_sources}
plot_google_map(data = gps_coordinates, 
                longitude = gps_coordinates$longitude, 
                latitude = gps_coordinates$latitude,
                longitude_limits = c(-11.5,38),
                latitude_limits = c(35,65),
                #shape_variable = gps_coordinates$country,
                fill_variable = gps_coordinates$source,
                center = c(10.08415861471605, # longitude; x-axis
                           49.905624997321596), # latitude; y-axis
                zoom_factor = 4, 
                map_source = "google")

ggsave(plot = last_plot(), filename = here("plots", "map.png"))

# zoom_factor: an integer from 3 (continent) to 21 (building), default value 10 (city). openstreetmaps limits a zoom of 18, and the limit on stamen maps depends on the maptype. "auto" automatically determines the zoom for bounding box specifications
```

The total daily volume of freshwater introduced into the system (`r totalV * ER` \si{\L\per\d}) was calculated based on the model system. Daily plant nutrient inputs via source water were calculated by gathering water analysis reports from locations all over Europe. Here, official reports provided by local authorities and water utilities in the corresponding municipalities were used to ensure that analyses were conducted according to accepted laboratory standards. The collected `r nrow(gps_coordinates)` water analysis reports originated from a total of `r length(unique(gps_coordinates$country))` countries, with their spatial distribution being shown in Figure \@ref(fig:map_water). 

\begin{figure}
\centering
  \includegraphics{../plots/map.png}
  \caption{Geographical distribution of water analysis reports included in the source water dataset. Green: analysis of municipal tap water; Red: rainwater analysis.}
  \label{fig:map_water}
\end{figure}

In water analysis, it is common practice to report the value of the instrument detection limit instead of the measured value in case that the analyte concentration found was below the sensitivity of the measurement instrument (ISO 11843, 2020). This practice is known as data censoring and used to ensure that the analyte signal can be distinguished from the instruments signal noise with satisfying confidence. However, retaining these data points in the dataset leads to incorrect results as the true concentration is within the interval from zero to the detection limit. To avoid an overestimation of plant nutrient concentrations within the tap water, censored data were recalculated by \gls{mle}, using the \emph{cenmle()} function from the R package NADA [@Helsel2011]. 



### Aquafeeds
The total daily feed input $m_{in,total}$ (`r rearingV * SD * (FR/100)` \si{\kg\per\d}) was calculated based on the model system. Eventually, daily nutrient inputs via aquafeeds were calculated by multiplying the total daily feed input with the inclusion rates of the corresponding nutrient $\gamma_{X}$. As the study focusses on plant nutrients in aquaponic systems, nutrient retention by the livestock was considered, because it renders nutrients inaccessible for plants. This was done by multiplication of the total daily nutrient input with an \gls{adc} for each nutrient, yielding the true daily plant nutrient input via aquafeed. It was assumed that \gls{cp} digestibility would be on average \SI{90}{\p}, while the \gls{adc} for all other plant nutrients were assumed to be \SI{50}{\p} [@Lall2002]. Finally, due to the fact that solid substances are not accessible for plant uptake, it was assumed that \SI{50}{\p} of the digestible fraction is retained while the remaining \SI{50}{\p} is excreted in dissolved form [@Halver2003]. The calculation is described by equation \ref{eq:feed_input}.

\begin{equation}
  m_{in,true} = m_{in,total} \cdot r_{feed} \cdot \gamma_{X} \cdot \text{ADC}_{X} \cdot r_{ex,diss}
  \label{eq:feed_input}
\end{equation}



### Alkalinity supplements
An estimate for the total daily input of alkalinity supplements was derived from the total daily feed input. First, the daily \gls{cp} input was calculated by multiplying the total daily feed input with the \gls{cp} inclusion rate of the feed. 
..




## Statistics
\SI{90}{\p} confidence intervals were calculated for the nutrient contribution of aquafeeds and tap water.

## Software
All calculations were conducted using R (v4.2.2) with RStudio (“Spotted Wakerobin” Release) as graphical user interface.




# Aquafeeds as nutrient sources

```{r}
source(here("R", "plots", "feed_plots.R"))
```


Aquafeeds are considered to be the most important source of nutrients in aquaculture and, consequently, aquaponics, providing a large and continuous nutrient input into the system. Formulating a specific aquaponics feed is thus generally thought to be the most suitable approach to develop an “off-the-shelf” nutrient delivery system [@Lennard2017; @Eck2019]. However, detailed information with regards to the inclusion levels of nutrients in aquafeeds is scarce.

In fish, the essentiality of calcium, phosphorus, magnesium, sodium, potassium, chloride, iron, copper, manganese, zinc, iodine, selenium, cobalt, and chromium was shown [@Lall2021]. These elements play a vital role in the development and maintenance of the skeletal system, signal transmission, osmoregulation, as reactive centers of metalloenzymes, and others. A comprehensive summary can be found can be found elsewhere (see e.g., Lall, 2003).
While some of the mentioned substances are taken up in relatively large quantities, only small amounts of others are required, as indicated by the term "trace element". However, the exact requirements are unknown for many of the above-mentioned nutrients due to difficulties in their determination, as described elsewhere (see e.g., Lall, 2003). In addition, uptake from the water via gills and skin contributes significantly to the total nutrient budget, enabling the animal to reach homeostasis. This is known for

Following: [@Lall2003]
K (water, feed)
Na (water)
Cl (water)
Following: [@Lall2021]
Ca (gills)
I (80% water rainbow trout)
Cr (gills)
Co (gills and feed)
Cu (feed; gills can contribute more than half of demand)
P (feed)
Mg (feed)
Fe (feed)
Se (feed)
Zn (feed)
Mn (unknown)

Therefore, the need to include those specific nutrients into the feed formulation is usually not done.

```{=tex}
% list of Bittzansky nutrients for fish and plants, together with those that have to be reported by law
```


With regards to commercial aquafeeds from producers located in Europe, this circumstance is due to the current regulations affecting the labeling. Currently, general mandatory and specific requirements with regards to the labelling of compound feeds are defined by Regulation (EC) No. 767/2009 of the European Parliament defines in the articles 15 and 17 and a detailed explanation with respect to the labelling of feed materials and compound feed for animals intended for human consumption can be found in Annex VI, stating that the levels of crude protein, crude fibre, crude oils and fats, crude ash, calcium, sodium and phosphorus need to be listed by the feed producer (EU). This 


## Nutrient inclusion rates


While it is common standard to report the proximate composition of the diets, giving Kjeldahl nitrogen and ash as sum parameter for the metal content of feeds, more detailed insights with respect to other nutrients essential for plant growth are rather scarce.


### Commercial feeds
Information about nutrient inclusion rates that exceed the standard fish nutrients is scarce as it is not of greater interest for aquaculture and the growth of the fish. A publication looking into this topic dates back 40 years [@Tacon1983]. 

```{=tex}
% overview table with the composition of commercial feeds
```


### Experimental feeds
Several experiments were conducted in the past to closer examine the extent to which manipulation of aquafeeds is possible when aiming at specific substance excretion patterns by fish.

@Seawright1998

Different natural and new feed ingredients were recently tested.
@Shaw2022a
@Shaw2022b
@Gebauer2023


```{=tex}
% overview table with the composition of experimental feeds
```

## Sources and extent of variability
[@Prabhu]

### Fish species
@Gebauer2023
@Shaw2022b

### Ingredients (digestibility)
@Shaw2022a
@Shaw2022b
@Gebauer2023
@Seawright1998


When discussing aquafeeds as nutrient input route, it is important to consider pathways of diet utilization, as shown in Figure 1.2. Uneaten feed and the indigestible mass fraction of ingested feed make up the solid wastes. The digestible mass fraction is meanwhile utilized to sustain the animals’ basal metabolism, somatic growth and reproductive activity. Metabolic end products are then excreted via the gills and the urinary system in form of dissolved matter [@Hardy2003; @Evans2005]. Only the dissolved fraction of the nutrients is immediately available for plant uptake via the root system.

The availability of digestibility, retentiton and excretion data for individual nutrients, with the former expressed as \gls{adc}, depends on whether the nutrient is essential and has to be considered for feed formulation. 

```{=tex}
% overview of ADCs determined for plant nutrients
```

N that is, for the most part, present as \gls{cp} in aquafeeds, is generally well-digestible, with an \gls{adc} usually being above \SI{70}{\p} and on average approximately \SI{90}{\p} [@Guillaume2001; @IAFFD2021]. The excretion of N as end product of the protein and amino acid catabolism takes place in form of ammonia (\ce{NH3}) and, to a small extent, urea. The predominant excretory site are the gills, followed by renal excretion [@Dabrowski2003]. 

A less digestible nutrient is P with \gls{adc} ranging from \SI{70}{\p} to only \SI{40}{\p} and a resulting excretion of \SI{30}{\p} to \SI{60}{\p} of the supply [@Lall2003; @Sugiura2018]. Especially plant ingredients in aquafeeds can cause low \gls{adc} if they are rich in phytic acid. Phytic acid is poorly digestible and can furthermore reduce the digestibility of minerals in the feed. This might also explain contradictory information in literature with reported renal excretion rates of \SI{90}{\p} of the total excreta [@Lall2003] in contrast to estimates of \SI{28}{\p} of excretion taking place in dissolved form and \SI{30}{\p} to \SI{64}{\p} excreted as particulate P [@Dabrowski2003]. 

Studies about \gls{adc} for the remaining essential plant nutrients are scarce. Variability of \gls{adc} among different feed ingredients was shown in Atlantic salmon (\emph{Salmo salar}) for Ca, Mg, Fe, Mn and Zn, with \gls{adc} ranging between \SIrange{30}{50}{\p} [@Sugiura1998]. Excretion of the earth alkaline metals Ca and Mg primarily occurs in dissolved form via the gills and urine [@Oikari1985; @Lall2003]. Mn, in contrast, is mostly excreted in solid form as feces, while renal excretion was found to be negligible [@Lall2003]. Cu is predominantly excreted via the bile [@Bury2003]. Excess dietary Cu is not taken up but excreted as feces. Cu inclusion rates in aquafeeds are thus reduced to minimize its release into the environment [@Lall2003]. Excretion of Zn mostly takes place renally and via the gills [@Lall2003]. 



# Water as nutrient source
- Why is water added?
- Where does it come from?


Water is added to the system either to compensate evaporation, spillage or leaning losses (e.g., the drum filter backwash) or to dilute accumulating substances such as nitrate [Maucieri]. The source water can originate from wells, collected rain or the municipal tap.

## Nutrient inclusion rates
The importance of water as nutrient source is seen as negligible [@Schmautz2016]. However, large amounts of calcium (Ca), magnesium (Mg) and sulfur (S), that are essential plant nutrients, were found to enter the system via exchange water in one study [@Delaide2017]. 

A recent survey revealed that the majority of European aquaponics entities (\SI{71.7}{\p}) uses municipal tap water [@Raulier2023]. Aquaponic food production is usually seen in the context of decentralized urban farming, where limited space due to high land prices is compensated by innovative approaches such as vertical farming and synergy effects due to the simultaneous production of fish and crops. Though, it has to be considered that the availability of rain water and space for its storage is limited and wells are unlikely to be approved by authorities. Thus, tap water would 

This observation will likely 

This is not surprising, considering that aquaponic facilities 


### Sources and extent of variability
Though, water can vary in its composition, depending on its origin. A glimpse of the extent of variability of terrestrial waters in their chemical composition is provided in Figure 1.1. Weathering of rock, ion exchange, redox reactions and the buildup of biomass are the main processes that are affecting concentrations of the shown compounds [@Stumm1981]. The highest variability can be found in anionic compounds such as nitrate (\ce{NO3-}) and sulfate (\ce{SO4^2-}) and the cationic alkaline earth metals \ce{Ca^2+} and \ce{Mg^2+} that are covering a concentration range of approximately two orders of magnitude.





# Alkalinity supplements as nutrient sources
The term alkalinity supplements summarizes several alkaline substances that are used to maintain a stable pH in the water [@Timmons2010]. Nitrification decreases the pH over time and, consequently, also the activity of nitrifyers [Ward2011]. Thus, a stable pH has to be maintained to ensure both high nitrification performance and animal welfare. For this purpose, several Na-based substances such as sodium hydrogen carbonate (baking soda, \ce{NaHCO3}) are commonly used in aquaculture due to their high and rapid solubility at a comparably cheap price. However, high sodium (Na) concentrations must be avoided in aquaponic systems due to its phytotoxicity [@Maathuis2014]. Na is therefore replaced with several other supplements based on K or Ca that come with the benefit of providing an additional source of nutrients besides increasing the pH.

## Nutrient inclusion rates


## Sources and extent of variability
Depends on the product quality




# The big picture


## Contribution of the nutrient sources to the total nutrient mass flows
The average nutrient composition of both aquafeeds and source water and the lower and upper limits of the \SI{90}{\p} confidence interval for source water are shown in Table 3.1. Figure 3.2 provides an additional graphical presentation of the findings with emphasis on the variability of nutrient contributions by source water.

The highest daily nutrient inputs with respect to the macronutrients N and P in the reviewed studies likely originated from aquafeeds, being on average 99.2% and 99.3%, respectively. The overall amounts of these two nutrients contributed by source water are negligible. Aquafeeds can also be assumed being the main source of the micronutrients Fe, Mn, Cu, Zn, and Mo, with an average contribution of 96.2%, 97.2%, 79.8%, 93.9% and 76.3%, respectively. Among these nutrients, Cu and Mo showed the highest variability with respect to the contribution of source water between locations (Cu 90% CI: 8%, 29.7%; Mo 90% CI: 14.7%, 29.3%), with up to almost a third of the daily inputs possibly entering the system via the water. Meanwhile, source water likely had a comparably high contribution to Mg and S. With 76.5% for Mg (90% CI: 72.1%, 79.7%) and 68.8% for S (90% CI: 58.9%, 74.8%), an average delivery of more than half of the daily input was found. Also, a considerable contribution to daily B (63.1%; 90% CI: 48.5%, 48.5%) and Ni (56.2%; 90% CI: 33.3%, 67.4%) inputs probably originated from the source water.

Inputs of the remaining two nutrients, K and Ca, were found to be heavily depending on the alkalinity supplement used. In absence of the alkalinity supplement, aquafeeds were found to have the greatest impact on K with a contribution of 81.16%, while a higher proportion of Ca (69.62%; 90% CI: 65.51%, 72.86%) would enter the system on average via the water, compared with an average of30.38% of feed contribution. The difference is more pronounced with respect to Na: 82.25% (90% CI: 73.11%, 86.76%) of the total input was calculated to originate from water while only 17.75% were found to enter the system via daily feeding. However, the calculated quantities of alkalinity supplements necessary to maintain a constant pH were found to dominate each input scenario, resulting in a contribution of more than 80% of the total input of K, Ca, or Na in any case. 

## Economic considerations
Table 3.2 summarizes some supplements, their properties and prices. Using one of these substances would result in the supplement contributing 98.7%, 86.5% or 96.3% of the total K, Ca or Na input, respectively. Regarding the costs for alkalinity supplements, Na based substances are the cheapest, with prices of between 0.01 and 0.03 EUR d-1, followed by Ca based supplements ranging in the same price class. The most expensive supplements to be used are those based on K. These substances would cause costs between 0.10 and 0.14 EUR d-1, thus being on average 6 times more expensive than Ca or Na based substances.




The contribution of different nutrient sources to the total daily nutrient inputs into an aquaponic system were compared based on assumptions derived from literature.
It could overall be confirmed by the current input scenario that aquafeeds serve as major nutrient input route for N, P, K, Fe, Mn, Cu, Zn, and Mo after consideration of digestibility and nutrient retention by fish. These results confirm prior findings [@Delaide2017]. Considering variability among locations, it could furthermore be confirmed that Ca, Mg, S, B and Na originate from source water. However, the variability found in the case of Ca does not reflect the usual concentration range of approximately two orders of magnitude found in terrestrial waters [@Stumm1981]. A study reviewing the Ca concentration in tap waters from the USA and Canada found a range from \SIrange{1}{135}{\mgL} [@Morr2006]. 

Considering that the obtained results were obtained making use of assumptions, these initial assumptions need to be discussed. The species distribution with strong emphasis on Tilapia (\emph{Oreochromis} spp.) and African catfish (\emph{Clarias gariepinus}) matches with survey results obtained in Europe [@Villarroel2016], even though approximately half of the studies included in the dataset were conducted in the USA. However, surveys conducted in the USA also revealed that the species mostly cultivated in aquaponic systems are Tilapia [@Love2015; @Pattillo2022]. On the other hand, the average stocking density of 7.69 kg m-3 that was used for the nutrient contribution calculations can neither be considered representative for the intensive cultivation of Tilapia nor for African catfish. While densities between \SIrange{10}{50}{\kg\per\cubicm} are reported as acceptable for cultivation in RAS systems [@El-Sayed2019], densities of \SI{120}{\kg\per\cubicm} were reported to be tolerated by the latter species without any adverse effects on fish health [@Nieuwegiessen2009]. Even though salmonids such as Rainbow trout (\emph{Oncorhynchus mykiss}) are seldom reared in aquaponic systems, they can tolerate stocking densities up to \SI{137}{\kg\per\cubicm}. Though, normal stocking densities range between \SIrange{10}{20}{\kg\per\cubicm}. The assumptions are rather on the lower end for salmonids [@Pennell1996]. Common carp (\emph{Cyprinus carpio}) is irrelevant for aquaponic systems due to its low market price. Though, it was found that fish welfare can be guaranteed at a stocking density of around \SI{28}{\kg\per\cubicm}, which is $\approx 4$ times higher than the assumed density [@Ruane2002]. Consequently, it can be expected that the contribution of aquafeeds to the total daily nutrient input under intensive rearing conditions for the mentioned fish species is considerably higher than assumed in this study. Though, the presented nutrient contribution scenario might be representative for the cultivation of Pikeperch due to the comparably low stocking densities ($\approx$\SI{5}{\kg\per\cubicm}) recommended for its cultivation [@Kestemont2015]. The average \gls{fr} of \SI{2}{\p} is in line with recommendations for the husbandry of Tilapia, Channel catfish (Ictalurus punctatus) and Common carp [@Lovell2003]. Another aspect with respect to the contribution of feeds is that the feed composition is, in adaptation to the species to be reared, variable. For instance, crude protein inclusion rates in commercial aquafeeds for the mentioned species range from 32% for O. spp. [@Wilson2003; @El-Sayed2019] to \SI{50}{\p} for Pikeperch [@Geay2015]. Further variation in the nutrient composition is introduced by feed manufacturers due to changes in its raw material composition made because of fluctuations in feed ingredient availability and market prices. While changes in the mass concentrations of crucial nutrients such as the \gls{cp} inclusion rate are acceptable only within tight limits, this might not be the case for other nutrients that are disregarded in the formulation process. Overall, data about the inclusion rates of nutrients relevant for plants in commercial aquafeeds can be considered scarce because these compounds are either not essential for aquatic animals or can be taken up from water via the gills in sufficient quantities. These nutrients are thus usually not considered during feed formulation [@Lall2003]. However, a comprehensive review about the range of plant nutrient mass concentrations in commercial aquafeeds has not yet been conducted to the best of the authors knowledge. The mean water exchange rate of 3% can be considered at the lower end of commonly reported water exchange rates for RAS and within the range recommended for aquaponic systems [@Timmons2010]. 

Using analysis reports from municipal water suppliers can be considered as valid approach with respect to the acquisition of nutrient concentration data in source water as it was found by a prior survey that educators, including research institutions, usually use tap water [@Love2015]. It could be further argued that, due to its low water demand, aquaponics is generally seen as food production system suitable for urban areas or arid regions [@Kloas2015; @Joyce2019]. Establishing an aquaponic system in these regions comes with limited access to water sources such as well water or rivers and lakes. Rainwater, on the other hand, would require large storage capacities. Tap water is thus assumed to be the most important water source. Recalculation of nutrient concentrations as response to high proportions of censored data where only the limit of detection is reported is a rather uncommon approach. However, removing the affected observations would have resulted in a loss of a large proportion of data and, in addition, would have led to an average value that is higher than the estimates obtained by recalculation because the remaining data consisted only of concentration values that are high enough to be determined. A similar result would have arisen from using the reported detection limits for the calculation of nutrient concentration means. The results of the nutrient concentration recalculation via MLE procedure can thus be considered closer to the true concentration means than the outcome of the other mentioned approaches. 

The daily quantity of alkalinity supplements to be added might be overestimated. In biofilters, both nitrification and denitrification processes usually occur due to partially anoxic conditions, for instance within biofilms. Denitrification yields alkalinity [@Timmons2010]. However, the contribution of alkalinity supplements will likely still exceed 50% of the total daily nutrient inputs. Furthermore, the calculations assumed that the supplements would have a purity of 100%. In practice a lower purity, corresponding to a food grade certification, can be expected.



## Implications for nutrient management in aquaponic systems
Firstly, considering potential precipitation, target nutrients for an increased delivery could include N, K, Mg, S, B, and Zn. This suggestion is partly confirmed by prior empirical studies. One of those found K, Mg, and Zn to accumulate out of the mentioned nutrients, with additional accumulation of P, Mn and Zn [@Seawright1998]. Another study reported N, K, and Mg accumulation alongside P and Cu (Shaw et al. 2022). The focus on K and Mg can thus be considered “safe”, considering the generally high solubility of both compounds [@Lide2007] and reported deficiencies for both in aquaponics [@Lunda2019]. Natural feed ingredients rich in K are soy lecithin (\SI{13.0}{\p} K on dry matter basis), vinasse (\SI{3.6}{\p} K on dry matter basis) and some protein feedstuffs such as poultry by-product meal (\SI{3.5}{\p} K on dry matter basis) and hydrolyzed fish solubles \SIrange{2.4}{2.9}{\p} K on dry matter basis). Mg can be introduced to aquafeeds mostly via the use of seaweed (\SI{3.9}{\p} Mg on dry matter basis) (International Aquaculture Feed Formulation Database (IAFFD) 2021). The drawback of the use of some of the above-stated feed ingredients is that the target nutrient is usually accompanied by comparably high amounts of Na. Even though this study shows that aquafeeds were not the main input route of Na in the reviewed literature, it is nevertheless reasonable to attempt to decrease Na inclusion rates during feed formulation because the initial assumptions leading to the nutrient input scenarios are not representative for commercial systems, as described above. This could be done for instance by replacement of feed ingredients with high Na inclusion rates such as fish meal originating from marine fish by other protein feedstuffs. 

Another possibility is given by using food-grade salts of the desired nutrients. However, considering the nutrient input via feed (\SI{1.1}{\p}) in comparison with that of alkalinity supplements containing K (\SI{98.7}{\p}) in this study, it depends on target concentrations and costs whether it is reasonable or not to increase the inclusion rate of K in a tailored aquafeed versus using the respective alkalinity supplements. 

It remains questionable whether it is more meaningful to include, for instance, salts into the feed or to add them directly into the system. Inclusion rates of nutrients that are confirmed to be affected by precipitation (Ca, Fe) are, in turn, recommended to be decreased in case the nutrient is not essential for the livestock and a change in feed formulation economically viable.


## Other implications


# Conclusion
This article provides a synthesis of information concerning the sources of nutrients in aquaculture and aquaponic systems, their contribution to the total daily nutrient mass flows and their variability.


# References {-}

