---
title: "Nutrient inputs into aquaculture systems: Sources, variability, contribution to the nutrient budget and implications for integrated food systems such as aquaponics"
author:
  - name: Anıl Axel Tellbüscher
    email: atellbuscher@frov.jcu.cz
    affiliation: USB
    footnote: 1
  - name: Radek Gebauer
    email: rgebauer@frov.jcu.cz
    affiliation: USB
    footnote: 2
  - name: Jan Mráz
    email: jmraz@frov.jcu.cz
    affiliation: USB
    correspondingauthor: true
address:
  - code: USB
    organization: University of South Bohemia in České Budějovice, Faculty of Fisheries and Protection of Waters, South Bohemian Research Center of Aquaculture and Biodiversity of Hydrocenoses, Institute of Aquaculture and Protection of Waters
    addressline: Na Sádkach 1780
    city: 37005 České Budějovice
    country: Czech Republic
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Understanding substance inputs into aquaculture systems and their variability is necessary for their downstream management. This study reviews substance input routes with respect to their contribution to total substance flows.
keywords: 
  - aquaponics
  - minerals
  - fish feed
  - alkalinity supplements
journal: "Reviews in Aquaculture"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: library_APO-SUB-2.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    includes:
      in_header: ../latex/preamble.tex
    keep_tex: true
    citation_package: default
---

```{r setup, include=FALSE}
# Knitr options
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
options(digits = 1)

# Load packages
library(here)
library(readxl)
library(tidyverse)
library(rstatix)
library(NADA)
library(ggmap)
library(gdata)
library(viridis)
library(RColorBrewer)
library(bookdown)
library(png)
library(tellbuschR) # Install from GitHub TellAnAx/tellbuschR with GitHub token for authentication

# Load functions
source(here("R", "plot_functions.R"))
```

# Introduction
One of the obstacles in the wider adoption of aquaponic systems, the combination of aquaculture and hydroponics (soilless plant production), is stated to be the rather complex management of plant nutrients within the system. While a wide range of "off-the-shelf" fertilizer is available for hydroponics, this is not the case for aquaponic systems. Developing more efficient and simple nutrient management practices thus appears to be an opportunity to facilitate systems management. Within this context, nutrient management refers to guaranteeing that optimum plant nutrient concentrations are met by controlling nutrient inputs.

In aquaculture and aquaponics, nutrients enter the system through the daily ration of feed for the aquatic livestock (hereinafter denoted as aquafeed), alkalinity supplements used for pH control, and the source water [@Rakocy2006; @Eck2019; @Robaina2019]. When focussing on the efficiency aspect, the most efficient solution for nutrient management would be to monitor each plant nutrient individually. This would ensure that each individual nutrient can be supplied in the necessary amount to reach target concentrations, for instance as pure salts. However, a full-spectrum monitoring of all plant nutrients is not practical, as it requires expensive equipment and qualified workforce. Instead, it would be more meaningful to focus on the simplicity aspect and develop a management practice that is easily implementable without introducing additional tasks to the working schedule. Against this background, optimizing an existing nutrient input can be considered as more feasible solution. Though, when optimizing one nutrient input while not controlling for the others, it is important to consider the range of nutrient input that can be expected by other sources of nutrients. Awareness about the nutrients that run the risk of under- or oversupply under "non-average" conditions is crucial, as this might have adverse effects on the livestock or plants. Hence, the variability of nutrients originating from different sources ultimately determines which nutrients to focus on when optimizing nutrient inputs. However,little is known about the variability of plant nutrient mass flows in aquaculture, which is crucial for the development of a nutrient management solution that is optimized with regards to its simplicity of use, such as a dedicated aquafeed. 

The objective of this study is to assess the nutrient budget of aquaculture systems with regards to the contribution and variability of aquafeeds, alkalinity supplements, and the source water. Emphasis is given to aquaponics as this knowledge is expected to contribute to the development of simple means of nutrient management in this food production system.





# Methodology

## Model system

```{r}
source(here("R", "assumptions.R"))
```

The ASTAF-PRO system [@Kloas2015] was chosen as a model system for the estimation of aquafeed, alkalinity supplement, and source water mass flows into an exemplary aquaponic system. The sizing of the chosen system is comparable with results from a recent survey targeting aquaponics production sites [@Pattillo2022] and its system components are well-described. All model data are summarized in table \ref{tab:assumptions}.

```{=tex}
\include{table_assumptions.tex}
```


## Data collection and calculations

### Source water

```{r import_data}
source(here("R", "import_data.R"))
```

```{r data_wrangling}
source(here("R", "data_wrangling.R"))
```

```{r plot_functions}
source(here("R", "plot_functions.R"))
```

```{r plot_water_sources}
plot_google_map(data = gps_coordinates, 
                longitude = gps_coordinates$longitude, 
                latitude = gps_coordinates$latitude,
                longitude_limits = c(-11.5,38),
                latitude_limits = c(35,65),
                #shape_variable = gps_coordinates$country,
                fill_variable = gps_coordinates$source,
                center = c(10.08415861471605, # longitude; x-axis
                           49.905624997321596), # latitude; y-axis
                zoom_factor = 4, 
                map_source = "google")

ggsave(plot = last_plot(), filename = here("plots", "map.png"))

# zoom_factor: an integer from 3 (continent) to 21 (building), default value 10 (city). openstreetmaps limits a zoom of 18, and the limit on stamen maps depends on the maptype. "auto" automatically determines the zoom for bounding box specifications
```

The total daily volume of freshwater introduced into the system (`r totalV * ER` \si{\L\per\d}) was calculated based on the model system. Daily plant nutrient inputs via source water were calculated by gathering water analysis reports from locations all over Europe. Here, official reports provided by local authorities and water utilities in the corresponding municipalities were used to ensure that analyses were conducted according to accepted laboratory standards. The collected `r nrow(gps_coordinates)` water analysis reports originated from a total of `r length(unique(gps_coordinates$country))` countries, with their spatial distribution being shown in Figure \ref{fig:map_water}. 

\begin{figure}
\centering
  \includegraphics{../plots/map.png}
  \caption{Geographical distribution of water analysis reports included in the source water dataset. Green: analysis of municipal tap water; Red: rainwater analysis.}
  \label{fig:map_water}
\end{figure}



### Aquafeeds
The total daily feed input $m_{in,total}$ (`r rearingV * SD * (FR/100)` \si{\kg\per\d}) was calculated based on the model system. Eventually, daily nutrient inputs via aquafeeds were calculated by multiplying the total daily feed input with the inclusion rates of the corresponding nutrient $\gamma_{X}$. As the study focusses on plant nutrients in aquaponic systems, nutrient retention by the livestock was considered, because it renders nutrients inaccessible for plants. This was done by multiplication of the total daily nutrient input with an \gls{adc} for each nutrient, yielding the true daily plant nutrient input via aquafeed. It was assumed that \gls{cp} digestibility would be on average \SI{90}{\p}, while the \gls{adc} for all other plant nutrients were assumed to be \SI{50}{\p} [@Lall2003]. Finally, due to the fact that solid substances are not accessible for plant uptake, it was assumed that \SI{50}{\p} of the digestible fraction is retained while the remaining \SI{50}{\p} is excreted in dissolved form [@Halver2003]. The calculation is described by equation \ref{eq:feed_input}.

\begin{equation}
  m_{in,true} = m_{in,total} \cdot r_{feed} \cdot \gamma_{X} \cdot \text{ADC}_{X} \cdot r_{ex,diss}
  \label{eq:feed_input}
\end{equation}



### Alkalinity supplements
The input of alkalinity supplements was estimated based on @Timmons2010, with an amount of \SI{250}{\g} sodium bicarbonate (\ce{NaHCO3}) being recommended per \si{\kg} of feed input. Given the molar mass of \ce{NaHCO3} of $M(\ce{NaHCO3}) = \SI{84.01}{\gmol}$, this would translate into $n = \frac{m}{M} = \frac{\SI{250}{\g}}{\SI{84.01}{\gmol}} = \SI{2.975}{\mol}$ of \ce{HCO3-} as $n(\ce{HCO3-}) = n(\ce{NaHCO3})$.  




## Data presentation, statistics, and software
Mass concentrations of molecules were converted to element concentrations.

For all statistical tests, a significance level of $\alpha = 0.05$ was chosen, unless otherwise stated.

In water analysis, it is common practice to report the value of the instrument detection limit instead of the measured value in case that the analyte concentration found was below the sensitivity of the measurement instrument [@ISO11843]. This practice is known as data censoring and used to ensure that the analyte signal can be distinguished from the instruments signal noise with satisfying confidence. However, retaining these data points in the dataset leads to incorrect results as the true concentration is within the interval from zero to the detection limit. To avoid an overestimation of plant nutrient concentrations within the tap water, true means from censored data were estimated by \gls{mle}, using the \emph{cenmle()} function from the R package NADA [@Helsel2011]. The calculated estimates were reported along with their \SI{95}{\p} confidence intervals.

All calculations were conducted using R (v4.2.2).




# Feed(-ing) as nutrient source
Aquafeeds are considered the most important nutrient input into aquaculture systems. This is acknowledged by the availability of special feeds for open and closed system aquaculture. The former is, for instance, pond or cage aquaculture, where phosphorus emissions into receiving water bodies are intended to be reduced to avoid eutrophication. The latter is represented by \gls{ras}. Here, special feeds target at improved faeces stability and thus removability by mechanical filtration. Aquaponics feeds, meanwhile, contain high proportions of phosphorus and potassium. The formulation of a specific aquaponics feed that meets the nutritional requirements of both fish and plants by leading to higher excretion of otherwise deficient plant nutrients is generally thought to be the most suitable approach for the development of an “off-the-shelf” nutrient delivery system that does not require additional management effort by aquaponics farmers [@Lennard2017; @Eck2019]. However, a problem that comes with this approach is that fish and plants differ in their nutritional requirements, as shown in Table \ref{tab:essentials}.

```{=tex}
% list of Bittzansky nutrients for fish and plants, together with those that have to be reported by law
\include{table_essential_nutrients.tex}
```

Both fish and plants require carbon and oxygen in large quantities. Further macronutrients that are required by both are calcium, phosphorus, magnesium, iron, copper, manganese, and zinc for 

These elements are required for physiological functions such as the development and maintenance of the skeletal system, processes such as signal transmission, osmoregulation, as reactive centers of metalloenzymes (e.g., Haemoglobin, Chlorophyll), and others [@Lall2021; @Marschner2012]. 

However, detailed information with regards to the aquafeed inclusion levels of those nutrients is scarce. This might be due to the fact that nutritional requirements of some of the mentioned substances have not yet been established for aquatic animals. Thus, there is no incentive to include those nutrients into feed formulation or to monitor their inclusion rates, making considerable variation between feeds likely. 

In addition, current European regulations affecting labelling of compound feeds for aquaculture (defined by Regulation EC No. 767/2009 of the European Parliament) list only crude protein, crude fibre, crude oils and fats, crude ash, calcium, sodium and phosphorus as mandatory values to be stated on labels of commercial feeds. There is thus no obligation to analyse metals other than the stated ones. As a result, the most recent publication looking into metal inclusion rates in aquafeeds dates back 40 years [@Tacon1983]. However, comparing modern feeds with the results of that time (see Table \ref{tab:feedcomp}), changes in the feed composition are barely observable. 

```{r data_commercial}
data_commercial <- data$feed_commercial[[1]] %>% 
  filter(!target_system %in% c("aquaponics", "pond", "cage")) %>%
  filter(feed_category3 != "broodstock") %>% 
  select(-c(producer, feed_name, reference_id, target_system))
  
data_commercial_summary <- data_commercial %>% 
  create_plot_data() %>% 
  select(c(target_species:feed_category4, name:value)) %>% 
  drop_na(value) %>% 
  group_by(name) %>% # or growth stage? feed_category3
  summarise(
    n = n(),
    min = min(value, na.rm = TRUE),
    q1 = quantile(value, na.rm = TRUE, 0.25),
    q2 = quantile(value, na.rm = TRUE, 0.5),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    cv = sd/mean*100,
    q3 = quantile(value, na.rm = TRUE, 0.75),
    max = max(value, na.rm = TRUE)
    ) %>% 
  arrange(cv) %>% 
  print()
```

- N data

After exclusion of specialty feeds, the range in which phosphorus, potassium, and calcium can be found in modern commercial fish feeds is comparably narrow, ranging from \SIrange{7}{18}{\gkg}, \SIrange{10}{13}{\gkg}, and \SIrange{12}{20}{\gkg}, respectively. These nutrients also show the lowest \gls{cv} (between \SIrange{20}{21}{\p}). The average inclusion rates did not significantly change in comparison with past values, being \SI{12(2.5)}{\gkg}, \SI{11(2.3)}{\gkg}, and \SI{18(3.7)}{\gkg} of phosphorus, calcium, and potassium in nowadays' feeds, respectively, while \SI{14.9(0.02)}{\gkg}, \SI{9.6(1.0)}{\gkg}, and \SI{20.1(2.7)}{\gkg} were reported before [@Tacon1983]. Magnesium inclusion rates are slightly more variable, ranging from \SIrange{1}{3}{\gkg} with an average of \SI{2(0.8)}{\gkg} (\SI{2.1(0.1)}{\gkg} in the past) and a \gls{cv} of \SI{36}{\p}. 

In comparison with past values, the highest increase in their inclusion rates was found for the microminerals. 

The iron content of aquafeeds increased by `r (1-183.5/209)*100`\si{\p}, with an average of \SI{209(237.1)}{\mgkg} in comparison with \SI{183.5(24.7)}{\mgkg} in the 90ies. Iron is also the nutrient with the highest variability in aquafeeds, indicated by the \gls{cv} of \SI{114}{\p} and a range from \SIrange{40}{544}{\mgkg}.

Nowadays, aquafeeds are richer in copper than in the past. The average inclusion rate increased by `r (1-15/20)*100`\si{\p}, from \SI{15(1.4)}{\mgkg} to \SI{20(18.5)}{\mgkg}. In the case of zinc, inclusion rates increased by `r (1-90/207)*100`\si{\p}, from \SI{90(7)}{\mgkg} to \SI{207(126.6)}{\mgkg}. The inclusion rates of both copper and zinc were again highly variable, with \gls{cv} of \SI{93}{\p} and \SI{61}{\p}, and ranges from \SIrange{5}{46}{\mgkg} and \SIrange{90}{384}{\mgkg}, respectively. 

```{=tex}
\begin{sidewaystable}
  \caption{Synthesis of the mineral composition of starter and grower feeds for freshwater fishes suitable for aquaponic systems. Data covers compositions of fingerling and grower trout feeds reported by Tacon (1983), a compilation of feed compositions for African catfish, Common carp, Pikeperch, Nile tilapia, and Trout, the composition of speciality feeds for aquaponic systems, and some experimental feeds used in aquaponics trials. N, P, K, Ca, Mg: \si{\gkg}. Fe, Mn, Cu, Zn: \si{\mgkg}.}
  \label{tab:feedcomp}
  \vspace{1em}
  
  \csvautotabular[separator=semicolon]{table_feedcomp.csv}
  
\end{sidewaystable}
```



```{r plot_commercial_boxplot, include=TRUE}
data$feed_commercial[[1]] %>% 
  create_plot_data() %>% 
  filter(!target_system %in% c("aquaponics", "pond", "cage")) %>%
  select(c(producer:feed_category4, name:value)) %>% 
  drop_na(value) %>% 
  #ggplot(aes(x = target_species, y = value)) + 
  ggplot(aes(y = value)) + 
  geom_boxplot(notch = TRUE) +
  facet_wrap(facets = vars(name), scales = "free") + 
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
  )
```


```{r plot_commercial_density, include=TRUE}
data$feed_commercial[[1]] %>% 
  create_plot_data() %>% 
  filter(!target_system %in% c("aquaponics", "pond", "cage")) %>%
  select(c(producer:feed_category4, name:value)) %>% 
  plot_feed_comp()

ggsave(here("plots", "feed_composition_commercial.png"))
```



```{r data_commercial_special}
data_commercial_AP <- data$feed_commercial[[1]] %>% 
  filter(target_system %in% c("aquaponics")) %>% 
  select(-c(producer, feed_name, reference_id, target_system))
  
data_commercial_AP_summary <- data_commercial_AP %>% 
  create_plot_data() %>% 
  select(c(target_species:feed_category4, name:value)) %>% 
  drop_na(value) %>% 
  group_by(name) %>% 
  summarise(
    n = n(),
    min = min(value, na.rm = TRUE),
    q1 = quantile(value, na.rm = TRUE, 0.25),
    q2 = quantile(value, na.rm = TRUE, 0.5),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    cv = sd/mean*100,
    q3 = quantile(value, na.rm = TRUE, 0.75),
    max = max(value, na.rm = TRUE)
    ) %>%
  arrange(sd) %>% 
  print()
```


## Aquaponics feeds
There are some commercial feeds available that are marketed as specifically formulated for aquaponic systems (n = 3). Though, an in-depth characterisation of these feeds was not done until now to the best of the author's knowledge. 

```{r}
data$feed_commercial[[1]] %>% 
  create_plot_data() %>% 
  filter(!target_system %in% c("pond", "cage")) %>% 
  select(c(producer:feed_category4, name:value)) %>% 
  drop_na(value) %>% 
  mutate(target_system = if_else(is.na(target_system),"not specified",target_system)) %>% 
  filter(name == "P_gkg") %>% 
  rstatix::t_test(value ~ target_system)
```

However, even though available data is limited, it can be seen that the phosphorus content of these aquafeeds is highly variable (\gls{cv} = \SI{76}{\p}). The average phosphorus content of \SI{22(17)}{\gkg}, meanwhile, is not significantly different from the unspecific diets (t-test: p = 0.4). Information with regards to potassium and calcium inclusion rates were reported only for one feed. Here, potassium and calcium inclusion rates were `r (1-20/11)*100`\si{\p} higher and `r (1-15/18)*100`\si{\p} lower than the average of unspecific diets, respectively. Other data with regards to these feeds is not available to the best of the author's knowledge.




```{r include=TRUE}
data$feed_commercial[[1]] %>% 
  create_plot_data() %>% 
  filter(target_system %in% c("aquaponics")) %>% 
  select(c(producer:feed_category4, name:value)) %>% 
  plot_feed_comp()

ggsave(here("plots", "feed_composition_commercial.png"))
```


## Experimental feeds
In the recent past, experiments were conducted to closer examine the manipulability of substance excretion patterns by fish in relation to alterations in the raw material composition of provided aquafeeds.

```{r data_experimental_studies}
data_experimental <- data$feed_experimental[[1]] %>% 
  select(-c(reference_id, treatment_id))

data_experimental_summary <- data_experimental %>%   
  create_plot_data() %>% 
  drop_na(value) %>% 
  group_by(name) %>% 
  summarise(
    n = n(),
    min = min(value, na.rm = TRUE),
    q1 = quantile(value, na.rm = TRUE, 0.25),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    cv = sd/mean*100,
    q2 = quantile(value, na.rm = TRUE, 0.5),
    q3 = quantile(value, na.rm = TRUE, 0.75),
    max = max(value, na.rm = TRUE)
    ) %>% 
  print()
```

The overall nutrient composition of the experimental diets is comparable with commercially available aquafeeds, with no significant differences in average nutrient inclusion levels. Potassium (t-test: p = 0.3), phosphorus (t-test: p = 0.1), calcium (t-test: p = 0.4), and magnesium (t-test: p = 0.7) had average inclusion rates of \SI{9(2.8)}{\gkg}, \SI{14(4.9)}{\gkg}, \SI{15(8.4)}{\gkg}, and \SI{2(0.7)}{\gkg}, respectively.
Also, no differences were found for iron (t-test: p = 0.4), copper (t-test: p = 0.7), and zinc (t-test: p = 0.2). The average inclusion rates were \SI{348(29)}{\mgkg}, \SI{16(2.9)}{\mgkg}, and \SI{93(35.8)}{\mgkg}, respectively.

```{r}
bind_rows(list(commercial = data_commercial, experimental = data_experimental), .id = "source") %>% 
  t_test(Mg_gkg ~ source)
```

```{r include=TRUE}
data$feed_experimental[[1]] %>% 
  create_plot_data() %>% 
  plot_feed_comp()

ggsave(here("plots", "feed_composition_commercial.png"))
```



## Sources and extent of variability
The variability found in the data can be traced back to three causative factors, that are 1) the nutritional requirements of the fish species for which the feed was formulated, 2) the current approach of feed formulation based on nutrient digestibility, and 3) variability of the raw material composition among batches. 

[@Prabhu2016]

### Fish species
@Gebauer2023
@Shaw2022
@Shaw2022a

### Feed formulation
Feed formulators need to react to fluctuations of market prices for raw materials and the natural variation in raw material composition. 


### Ingredients 
@Shaw2022
@Shaw2022a
@Gebauer2023
@Seawright1998

 





Following: [@Lall2003]
K (water, feed)
Na (water)
Cl (water)
Following: [@Lall2021]
Ca (gills)
I (80% water rainbow trout)
Cr (gills)
Co (gills and feed)
Cu (feed; gills can contribute more than half of demand)
P (feed)
Mg (feed)
Fe (feed)
Se (feed)
Zn (feed)
Mn (unknown)



## Livestock as causative factor for variance
When discussing aquafeeds as nutrient input route, digestibility and retention of nutrients by the livestock must be considered. 

The indigestible mass fraction of ingested feed, together with uneaten feed, make up the solid wastes. The fraction of feed that is digestible is meanwhile utilised to sustain the animals’ basal metabolism, somatic growth and reproductive activity. Metabolic end products are then excreted via the gills and the urinary system in dissolved form [@Hardy2003; @Evans2005]. Only the dissolved fraction of the nutrients is immediately available for plant uptake via the root system.

The availability of \gls{adc} and nutrient retention data for individual nutrients depends on whether the nutrient is essential and has to be considered for feed formulation. 

```{=tex}
% overview of ADCs determined for plant nutrients
```

Nitrogen that is, for the most part, present as \gls{cp} in aquafeeds, is generally well-digestible, with an \gls{adc} usually being above \SI{70}{\p} and on average approximately \SI{90}{\p} [@Guillaume2001; @IAFFD2021]. The excretion of N as end product of the protein and amino acid catabolism takes place in form of ammonia (\ce{NH3}) and, to a small extent, urea. The predominant excretory site are the gills, followed by renal excretion [@Dabrowski2003]. 

A less digestible nutrient is phosphorus, with \gls{adc} ranging from \SI{70}{\p} to only \SI{40}{\p} and a resulting excretion of \SI{30}{\p} to \SI{60}{\p} of the supply [@Lall2003; @Sugiura2018]. Especially plant ingredients in aquafeeds can cause low \gls{adc} if they are rich in phytic acid. Phytic acid is poorly digestible and can furthermore reduce the digestibility of minerals in the feed. This might also explain contradictory information in literature with reported renal excretion rates of \SI{90}{\p} of the total excreta [@Lall2003] in contrast to estimates of \SI{28}{\p} of excretion taking place in dissolved form and \SI{30}{\p} to \SI{64}{\p} excreted as particulate phosphorus [@Dabrowski2003].

Studies about \gls{adc} for the remaining essential plant nutrients are scarce. Variability of \gls{adc} among different feed ingredients was shown in Atlantic salmon (\emph{Salmo salar}) for Ca, Mg, Fe, Mn and Zn, with \gls{adc} ranging between \SIrange{30}{50}{\p} [@Sugiura1998]. Excretion of the earth alkaline metals calcium and magnesium primarily occurs in dissolved form via the gills and urine [@Oikari1985; @Lall2003]. Manganese, in contrast, is mostly excreted in solid form with feces, while renal excretion was found to be negligible [@Lall2003]. Copper is predominantly excreted via the bile [@Bury2003]. Excess dietary copper is not taken up but excreted with feces. Copper inclusion rates in aquafeeds are thus reduced to minimise its release into the environment [@Lall2003]. Excretion of Zinc mostly takes place renally and via the gills [@Lall2003]. 




# Water as nutrient source
Aquaponics scholars usually focus on aquafeeds as the main source of nutrients, while the importance of water as nutrient source is denoted as negligible [@Schmautz2016]. However, a comparison of substance flows and their variability is lacking.

After the initial system fill-up, water is added to the system either to compensate evaporation, spillage or leaning losses (e.g., the drum filter backwash) or to dilute accumulating substances such as nitrate [@Maucieri2018]. The source water for an aquaponic system can originate from wells, collected rain, other sources of surface water, or municipal tap. However, considering that aquaponic systems are seen as especially beneficial for water-scarce regions [@Joyce2019] or urban areas [@Junge2017], where restrictive water management by authorities can be expected, tap water is likely the most important water source. Recent surveys corroborate this assumption, as the results show that the proportion of aquaponic practicioners using municipal tap water in the USA and Europe is \SI{47}{\p} and \SI{71.7}{\p}, respectively [@Pattillo2022; @Raulier2023]. Thus, the following assessment of water composition data focusses on the evaluation of tap water analyses. 


```{r}
data_water_summary <- data$water_tap[[1]] %>% 
  filter(Source == "tap") %>% 
  select(-c(Reference_ID:`EC_uS_cm_25degC`)) %>% 
  mutate(
    NO3_mgL = NO3_mgL*0.23 + NO2_mgL*0.3 + NH4_mgL*0.7, # convert to elemental N (TN_dissolved)
    SO4_mgL = SO4_mgL*0.3, # convert SO4 to elemental S
    PO4_mgL = PO4_mgL*0.33 # convert PO4 to elemental P
  ) %>% 
  pivot_longer(
    everything(),
  names_to = c("analyte", ".value"),
  names_pattern = "(.+)_(.+)*"
    ) %>% 
  mutate(
    analyte = as.factor(analyte),
    mgL = mgL * 1000,
    belowLimit = as.logical(belowLimit)
    ) %>% 
  drop_na() %>% 
  group_by(analyte) %>% 
  summarise(
    n = n(),
    min = min(mgL, na.rm = TRUE),
    lowconf = mean(cenmle(mgL, belowLimit, dist = 'gaussian'))[c(3)],
    meanconf = mean(cenmle(mgL, belowLimit, dist = 'gaussian'))[c(1)],
    sd = sd(mgL),
    cv = sd/meanconf*100,
    highconf = mean(cenmle(mgL, belowLimit, dist = 'gaussian'))[c(4)],
    max = max(mgL, na.rm = TRUE)
    ) %>% 
  mutate(
    analyte = if_else(analyte == "NO3", "N", analyte), # rename after conversion
    analyte = if_else(analyte == "SO4", "S", analyte), # rename after conversion
    analyte = if_else(analyte == "PO4", "P", analyte) # rename after conversion    
  ) %>% 
  #mutate(across(where(is.numeric), ~ num(., digits = 2))) %>% 
  arrange(desc(meanconf), cv) %>% 
  print() %>% 
  write_delim(here("interm", "summary_water.csv"), delim = ";")
```

```{=tex}
\begin{table}
\centering
  \caption{Concentrations of plant nutrients and other substances relevant for plant production in tap water samples from Europe. Thresholds given by the European Water Directive.}
  \label{tab:watercomp}
  \vspace{1em}
  
  \csvautotabular[separator=semicolon]{table_summary_water_edited.csv}
  
\end{table}
```

When reviewing analysis reports of water utilities (see Table \ref{tab:watercomp}), it can be seen that tap water is rich in the macronutrients calcium, sulfur, and magnesium. In accordance with this finding is that large amounts of calcium, magnesium, and sulfur, being essential plant nutrients, were found to enter the system via exchange water [@Delaide2017]. 
followed by potassium, nitrogen, phosphorus, followed by boron, iron, copper, zinc, and manganese. Interestingly, also the variability in their concentrations appears to follow this trend. 

Calcium, sulfur, and manganese were found to be comparably little variable among sampling sites in Europe, indicated by their \gls{cv} of \SIcv{53}, \SIcv{95}, and \SIcv{64}, respectively. Calcium concentrations ranged from \SIrange{3.3}{147.8}{\mgL}, with an average concentration of \SIci{59.9}{\mgL}{53.9}{66.0}. The second most abundant macronutrient was sulfur, ranging between \SIrange{0.4}{61.8}{\mgL}. On average, \SIci{12.5}{\mgL}{10.4}{14.8} sulfur is present in the water. The concentration range of magnesium in tap water was found to be \SIrange{0.1}{4.0}{\mgL}, averaging \SIci{12.0}{\mgL}{10.6}{13.5}. 

Other substances found in concentrations of \si{\mgL} are potassium, with a range of \SIrange{0.1}{23.6}{\mgL} and average of \SIci{2.76}{\mgL}{2.16}{3.36}, and nitrogen, ranging from \SIrange{0.09}{10.3}{\mgL} and averaging \SIci{2.16}{\mgL}{1.68}{2.65}. The most important nitrogenous compound in water is nitrate, making up the largest proportion of the total nitrogen in solution.

As far as other substances relevant for plant nutrition are analysed by water utilities, their concentrations in water are usually low due to their low solubility. Phosphorus in aqueous solution is mostly present in form of orthophosphates. The concentration of phosphorus as important plant nutrient is generally low in tap water. In the reviewed analysis reports, concentrations ranged from \SIrange{0.1}{6006}{\ugL}, averaging \SIci{223}{\ugL}{-67.2}{514}. The average concentrations of boron, iron, copper, zinc, and manganese were found to be similar. An average of \SIci{47}{\ugL}{30.3}{63.8} boron, \SIci{25.3}{\ugL}{3.35}{47.2} iron, \SIci{18.8}{\ugL}{3.97}{33.7} copper, \SIci{11.5}{\ugL}{1.17}{21.8} zinc, and \SIci{4.39}{\ugL}{2.57}{6.21} manganese was found, respectively. Among these microminerals, iron and zinc had the highest variability (iron: \gls{cv} = \SIcv{480}; zinc: \gls{cv} = \SIcv{448}), ranging from \SIrange{0.005}{1320}{\ugL} and \SIrange{0.5}{200}{\ugL}, respectively. The variability of manganese and boron was found to be lower (manganese: \gls{cv} = \SIcv{224}; boron: \gls{cv} = \SIcv{193}), with a mass concentration range of \SIrange{0.2}{77.1}{\ugL} and \SIrange{5}{638}{\ugL}, respectively.
Mo
Ni
Na
Cl



## Sources and extent of variability
Water can vary both spatially and temporally in its composition, depending on where it is sourced. Tap water can originate from natural springs, ground-, or surface water, where it was in contact with solids or gases for a prolonged period of time. 

The spatial variation of water is caused, for instance, by the properties of the aquifer. An important process that affects the composition of spring- and groundwater is weathering of rock, ion exchange, and the buildup of biomass [@Stumm1981]. In regions where water originates from surface water such as rivers or lakes, gas exchange and resulting chemical reactions, for instance redox reactions with oxygen, are affecting the composition of the water.

Temporal variation can be caused by weather events such as heavy rain, snow melting, floods, or anthropogenic inputs such as agricultural activity.

However, raw water is usually sourced 
might be treated by water utilities due to legislative reasons. 

In the European Union, the composition of tap water is regulated by the EU Drinking Water Directive [@EU2020]. A set of threshold values (see Table \ref{tab:watercomp}) ensures that the water is safe for lifelong human consumption and fulfills technical requirements.
The chemical parameters stated in Annex I, Part B, of the EU Drinking Water Directive are related to potential hazards and comprise threshold values for nitrate, nitrite, boron, copper, and nickel. In addition, technical indicator parameters are defined in Annex I, Part C. Those values affect the pH, conductivity, ammonium, iron, manganese, sulfur (sulphate), sodium, and chloride levels of the water. Comparing these While maximum (pH: also minimum) levels are thus given


A glimpse of the extent of variability of terrestrial waters in their chemical composition is provided in Figure 1.1. 




# Alkalinity supplements as nutrient sources
The term alkalinity supplements summarizes several alkaline substances that are used to maintain a stable pH in the water [@Timmons2010]. Nitrification decreases the pH over time and, consequently, also the activity of nitrifyers [Ward2011]. Thus, a stable pH has to be maintained to ensure both high nitrification performance and animal welfare. For this purpose, several Na-based substances such as sodium hydrogen carbonate (baking soda, \ce{NaHCO3}) are commonly used in aquaculture due to their high and rapid solubility at a comparably cheap price. However, high sodium (Na) concentrations must be avoided in aquaponic systems due to its phytotoxicity [@Maathuis2014]. Na is therefore replaced with several other supplements based on K or Ca that come with the benefit of providing an additional source of nutrients besides increasing the pH.

## Nutrient inclusion rates
- high in single nutrient (>90% purity)

## Sources and extent of variability
- variability very low; depends on the product quality




# Drawing the big picture


## Contribution of the nutrient sources to the total nutrient mass flows


Inputs of the remaining two nutrients, K and Ca, were found to be heavily depending on the alkalinity supplement used. In absence of the alkalinity supplement, aquafeeds were found to have the greatest impact on K with a contribution of 81.16%, while a higher proportion of Ca (69.62%; 90% CI: 65.51%, 72.86%) would enter the system on average via the water, compared with an average of30.38% of feed contribution. The difference is more pronounced with respect to Na: 82.25% (90% CI: 73.11%, 86.76%) of the total input was calculated to originate from water while only 17.75% were found to enter the system via daily feeding. However, the calculated quantities of alkalinity supplements necessary to maintain a constant pH were found to dominate each input scenario, resulting in a contribution of more than 80% of the total input of K, Ca, or Na in any case. 


## Stoichiometry of nutrient inputs



## Economic considerations





The contribution of different nutrient sources to the total daily nutrient inputs into an aquaponic system were compared based on assumptions derived from literature.
It could overall be confirmed by the current input scenario that aquafeeds serve as major nutrient input route for N, P, K, Fe, Mn, Cu, Zn, and Mo after consideration of digestibility and nutrient retention by fish. These results confirm prior findings [@Delaide2017]. Considering variability among locations, it could furthermore be confirmed that Ca, Mg, S, B and Na originate from source water. However, the variability found in the case of Ca does not reflect the usual concentration range of approximately two orders of magnitude found in terrestrial waters [@Stumm1981]. A study reviewing the Ca concentration in tap waters from the USA and Canada found a range from \SIrange{1}{135}{\mgL} [@Morr2006]. 

Considering that the obtained results were obtained making use of assumptions, these initial assumptions need to be discussed. The species distribution with strong emphasis on Tilapia (\emph{Oreochromis} spp.) and African catfish (\emph{Clarias gariepinus}) matches with survey results obtained in Europe [@Villarroel2016], even though approximately half of the studies included in the dataset were conducted in the USA. However, surveys conducted in the USA also revealed that the species mostly cultivated in aquaponic systems are Tilapia [@Love2015; @Pattillo2022]. On the other hand, the average stocking density of 7.69 kg m-3 that was used for the nutrient contribution calculations can neither be considered representative for the intensive cultivation of Tilapia nor for African catfish. While densities between \SIrange{10}{50}{\kg\per\cubicm} are reported as acceptable for cultivation in RAS systems [@El-Sayed2019], densities of \SI{120}{\kg\per\cubicm} were reported to be tolerated by the latter species without any adverse effects on fish health [@Nieuwegiessen2009]. Even though salmonids such as Rainbow trout (\emph{Oncorhynchus mykiss}) are seldom reared in aquaponic systems, they can tolerate stocking densities up to \SI{137}{\kg\per\cubicm}. Though, normal stocking densities range between \SIrange{10}{20}{\kg\per\cubicm}. The assumptions are rather on the lower end for salmonids [@Pennell1996]. Common carp (\emph{Cyprinus carpio}) is irrelevant for aquaponic systems due to its low market price. Though, it was found that fish welfare can be guaranteed at a stocking density of around \SI{28}{\kg\per\cubicm}, which is $\approx 4$ times higher than the assumed density [@Ruane2002]. Consequently, it can be expected that the contribution of aquafeeds to the total daily nutrient input under intensive rearing conditions for the mentioned fish species is considerably higher than assumed in this study. Though, the presented nutrient contribution scenario might be representative for the cultivation of Pikeperch due to the comparably low stocking densities ($\approx$\SI{5}{\kg\per\cubicm}) recommended for its cultivation [@Kestemont2015]. The average \gls{fr} of \SI{2}{\p} is in line with recommendations for the husbandry of Tilapia, Channel catfish (Ictalurus punctatus) and Common carp [@Lovell2003]. Another aspect with respect to the contribution of feeds is that the feed composition is, in adaptation to the species to be reared, variable. For instance, crude protein inclusion rates in commercial aquafeeds for the mentioned species range from 32% for O. spp. [@Wilson2003; @El-Sayed2019] to \SI{50}{\p} for Pikeperch [@Geay2015]. Further variation in the nutrient composition is introduced by feed manufacturers due to changes in its raw material composition made because of fluctuations in feed ingredient availability and market prices. While changes in the mass concentrations of crucial nutrients such as the \gls{cp} inclusion rate are acceptable only within tight limits, this might not be the case for other nutrients that are disregarded in the formulation process. Overall, data about the inclusion rates of nutrients relevant for plants in commercial aquafeeds can be considered scarce because these compounds are either not essential for aquatic animals or can be taken up from water via the gills in sufficient quantities. These nutrients are thus usually not considered during feed formulation [@Lall2003]. However, a comprehensive review about the range of plant nutrient mass concentrations in commercial aquafeeds has not yet been conducted to the best of the authors knowledge. The mean water exchange rate of 3% can be considered at the lower end of commonly reported water exchange rates for RAS and within the range recommended for aquaponic systems [@Timmons2010]. 




## Implications for nutrient management in aquaponic systems



## Other implications


# Conclusion
This article provides a synthesis of information concerning the sources of nutrients in aquaculture and aquaponic systems, their contribution to the total daily nutrient mass flows and their variability.


# References {-}

